---
title: "Probability And Statistical Inference Math 9102(TU059)"
author: "Mina Jamshidian"
output: html_notebook
Student Number: D20124995
---

For using this project the real data from Portuguese Secondary School Student Performance was used.
* Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7. (https://repositorium.sdum.uminho.pt/bitstream/1822/8024/1/student.pdf).

The Dataset:
this datasets would be used in this project.
* [sperformance-dataset.csv](https://github.com/minajm/Satistics-Interface/blob/main/sperformance-dataset.csv).


############# PREPARE ############

## Introduction:

For the preparation part of the assignment, all appropriate statistical concept for analyzing part would be used, which is consist of: 
 
Populations and samples 
Describing the sample
Hypothesis test aims to achieve.
Understanding of the issues
Describe the variables 


## Populations and samples

Generalizing explanations around total collections of items is the goal of statistical analysis.
The population is the entire items. Access to the whole populations is not possible so the small part of the population which is sample would be used for statistical analysis. The sample would give information about the whole population.
The dataset that we want to use it for this project is the sample populations about two secondary schools in the Alentejo region of Portugal.
This sample has 33 variables and 382 records which were collected from school and questions that students answered. This sample is about the performance of two majors (Math and Portuguese language).

## Describing the sample

The variables that exist in the sample populations with their descriptions would be explained in below.
For this part, the sample would be described by the summary statistics about the variables.



```{r, warning=FALSE}
#### Loading Packages

# Loading packages
library(pander)
library(pastecs)
library(tidyverse)
library(data.table)
library(cowplot)
library(summarytools)
library(ggplot2)
library(Hmisc)
library(psych)
library(gridExtra)
library(purrr)
library(gplots)
library(lsr)

```

```{r, warning=FALSE}
#clear workspace
rm(list = ls()) 
#set wd
setwd("~/Desktop/TU Dublin/SEM 1/Prob. and Statistical Inference/TEST2/CA-SI-1")
#check wd has been changed, make sure file is here
getwd() 

# Importing data
student_data <- read.csv("sperformance-dataset.csv",header=TRUE, stringsAsFactors=TRUE)

```

## Variables in the dataset
The variables that have been used in the dataset are listed below

```{r,warning=FALSE}

# Creating Table
panderOptions('table.split.table',Inf)
data_variables<-"
Variable Name           | Description
school                  | student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira)
sex                     | student's sex (binary: 'F' - female or 'M' - male)
age                     | student's age (numeric: from 15 to 22)
address                 | student's home address type (binary: 'U' - urban or 'R' - rural)
famsize                 | family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3)               
Pstatus                 | parent's cohabitation status (binary: 'T' - living together or 'A' - apart)
Medu                    | mother's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
Fedu                    | father's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
Mjob                    | mother's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
Fjob                    | father's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
reason                  | reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other')
guardian                | student's guardian (nominal: 'mother', 'father' or 'other')
traveltime              | home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)
studytime               | weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)
failures                | number of past class failures (numeric: n if 1<=n<3, else 4)
schoolsup               | extra educational support (binary: yes or no)
famsup                  | family educational support (binary: yes or no)
paid                    | extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)
activities              | extra-curricular activities (binary: yes or no)
nursery                 | attended nursery school (binary: yes or no)
higher                  | wants to take higher education (binary: yes or no)
internet                | Internet access at home (binary: yes or no)
romantic                | with a romantic relationship (binary: yes or no)
famrel                  | quality of family relationships (numeric: from 1 - very bad to 5 - excellent)
freetime                | free time after school (numeric: from 1 - very low to 5 - very high)
goout                   | going out with friends (numeric: from 1 - very low to 5 - very high)
Dalc                    | workday alcohol consumption (numeric: from 1 - very low to 5 - very high)
Walc                    | weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)
health                  | current health status (numeric: from 1 - very bad to 5 - very good)
absences                | number of school absences (numeric: from 0 to 93)
G1                      | first period grade (numeric: from 0 to 20)
G2                      | second period grade (numeric: from 0 to 20)
G3                      | final grade (numeric: from 0 to 20, output target)
"

df<-read.delim(textConnection(data_variables),header = FALSE,sep = '|',strip.white = TRUE,stringsAsFactors = FALSE)
names(df)<-unname(as.list(df[1,]))
df<-df[-1,]
row.names(df)<-NULL
pander(df,style='rmarkdown')

```
# Research Questions and Hypothesis
The Goal of this research is finding the  relationship between student grade in two majors (Math and Portuguese) at secondary school and several aspects of their life.

Research Questions:
The student final grade can be predicted by their study time and average of previous grades?

For finding the answer to the main question and getting good information about the data we have to answer the questions.

1.What is the average of previous student's grade?
2.What is the frequency distribution of study time?

3.What is the relationship between Final grades and average of previous grades?

Hypothesis 1
H0: there is no relationship between students Average previous grades and final grade for Portuguese major.
H1: there is a relationship between students Average previous grades and final grade for Portuguese major.

Hypothesis 2
H0: there is no relationship between students Average previous grades and final grade for Math major.
H1: there is a relationship between students Average previous grades and final grade for Math major.

4.What is the relationship between study time and final grade?

Hypothesis 3
H0: There is no difference in Final Portuguese Grade of students  and their study time.
HA: There is a difference in Final Portuguese Grade of students and their study time.

Hypothesis 4
H0: There is no difference in Final Math Grade of students and their study time.
HA: There is a difference in Final Math Grade of students and their study time.


Variables of interest
The variables of interest used in this research are shown below:

```{r, warning=FALSE}
# Create table
panderOptions('table.split.table',Inf)
data_variable2<-"
Concept                     | Variable Name | Statistical Type | Possible Values
first period grade          |  G1           |   Numeric        | Range from 0 to 20
second period grade         |  G2           |   Numeric        | Range from 0 to 20
final grade                 |  G3           |   Numeric        | Range from 0 to 20
weekly study time           |  studytime    |   Categorical    | 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour
"


df<-read.delim(textConnection(data_variable2),header = FALSE,sep = '|',strip.white = TRUE,stringsAsFactors = FALSE)
names(df)<-unname(as.list(df[1,]))
df<-df[-1,]
row.names(df)<-NULL
pander(df,style='rmarkdown')

```
## Understanding Of The Issues

In order to use this dataset, two features of the dataset must be met. First of all, the dataset should be large enough and secondly, it should be representative of the real characteristics. The sample dataset is not sufficient to be used as a sample of the general population of students attending Portuguese Secondary Schools. Two main features required for the sample to be sufficient are the size of the sample and whether or not it is representative. 

The size of the sample is important to make sure both the sample and real population have the same features that we are looking for. Also, the dataset must be representative because it should have and show the same exact behavior or characteristics of the actual population. 

It is important to understand that the purpose of the grades in the dataset is to understand student capability and core skills of the subject but this data is hard to extract because some students may have had health issues on the day of the exam. It is important to consider this fact can introduce bias.

Although 37 questions were asked and reviewed prior to the test, there could be some variables which have not been accounted for. The lack of sufficient variables can also impact the accuracy and correctness of our analysis.

Furthermore, the exam was taken on paper documents rather than using a software. This may also introduce some errors to the scoring system because some papers may have had errors during examination or they could have not been reviewed correctly (human errors)


## Quantitative Variable:

In this sample population, there are three numerical quantitative (grades for two subject, age and absences). For these variables, the centre point(Mean and median) and the spread (Min, Max, IQR and Range) of data should be described. From this statistical analysis we would understand the distribution and the shape of variables therefore it would be helpful for finding what type of tests could be used for hypothesis testing.
Just this quantitative variable from the sample population would be used for this project: grades ( pG1,pG2,pG3,mG1,mG2,mG3)
Finding the skew, kurtosis and Interquartile range (IQR) for quantitative variables would be helpful to find the outliers, clusters and gaps. 

Quantitative variables description for this dataset:
```{r warning=FALSE, fig.width=10, fig.align='center'}
#  Selecting just the quantitative variables from dataset
q_var <- student_data %>% select(age,absences.p,absences.m,pG1,pG2,pG3,mG1,mG2,mG3)
q_var

describe(q_var,IQR = TRUE,omit = TRUE)

# Distribution of numeric variable with visualization
g1<-ggplot(q_var,aes(x=age))+labs(x='Age', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$age,na.rm = TRUE),sd=sd(q_var$age,na.rm = TRUE)))+
  ggtitle('Distribution Age')
g1
g2<-ggplot(q_var,aes(x=absences.p))+labs(x='Absence Portuguese', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$absences.p,na.rm = TRUE),sd=sd(q_var$absences.p,na.rm = TRUE)))+
  ggtitle('Distribution Absence Portuguese')
g2
g3<-ggplot(q_var,aes(x=absences.m))+labs(x='Absence Math', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$absences.m,na.rm = TRUE),sd=sd(q_var$absences.m,na.rm = TRUE)))+
  ggtitle('Distribution Absence Math')
g3
g4<-ggplot(q_var,aes(x=pG1))+labs(x='PG1', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG1,na.rm = TRUE),sd=sd(q_var$pG1,na.rm = TRUE)))+
  ggtitle('Distribution PG1')
g4
g5<-ggplot(q_var,aes(x=pG2))+labs(x='PG2', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG2,na.rm = TRUE),sd=sd(q_var$pG2,na.rm = TRUE)))+
  ggtitle('Distribution PG2')
g5
g6<-ggplot(q_var,aes(x=pG3))+labs(x='PG3', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG3,na.rm = TRUE),sd=sd(q_var$pG3,na.rm = TRUE)))+
  ggtitle('Distribution PG3')
g6
g7<-ggplot(q_var,aes(x=mG1))+labs(x='MG1', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG1,na.rm = TRUE),sd=sd(q_var$mG1,na.rm = TRUE)))+
  ggtitle('Distribution MG1')
g7
g8<-ggplot(q_var,aes(x=mG2))+labs(x='MG2', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG2,na.rm = TRUE),sd=sd(q_var$mG2,na.rm = TRUE)))+
  ggtitle('Distribution MG2')
g8
g9<-ggplot(q_var,aes(x=mG3))+labs(x='MG3', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG3,na.rm = TRUE),sd=sd(q_var$mG3,na.rm = TRUE)))+
  ggtitle('Distribution MG3')
g9
grid.arrange(g1,g2,g3,g4,g5,g6,g7,g8,g9,nrow=3)
```
## Categorical Variables:

In this dataset, 28 categorical (qualitative) variables exist that they would be described the dataset. They would help us in dividing our sample on the essential characteristics of the sample set. Most of the categorical variable in this dataset is nominal by number.
The numerical category variables would not be like the quantitative variable because the number value for them just have the meaning code. So this type of variables could be described by frequently and terms of values.
Just these qualitative variables from the sample population would be used for this project: studytime, famsize, goout, Dalc, Fedu, Medu.
The statistics summarization for these variables would be described in below.

*Important* Just the variables which were not repeated in the dataset have been chosen.

Categorical variables description for this dataset:
```{r, warning=FALSE, fig.width=10, fig.align='center'}
#  Selecting just the categorical variables from dataset that do not repeated 
c_var <- student_data %>% select(school,sex,address,famsize,Pstatus,Medu,Fedu,
                                 Mjob,Fjob,reason,nursery,internet,guardian.p,
                                 traveltime.p,studytime.p,failures.p,schoolsup.p,
                                 famsup.p,paid.p,activities.p,higher.p,romantic.p,
                                 famrel.p,freetime.p,goout.p,Dalc.p,Walc.p,health.p)
# Statistics Summarization
freq(c_var)

# Statistics Summarization by visualization
c_var %>%
  gather() %>%
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_bar(colour = "black", aes(y = ..prop.., group =1, fill = ..count..))+
    scale_fill_gradient("Count",low="#132B43", high="#56B1F7")


```


############# EXplore And Model ############

## Introduction:
For the exploration part of the assignment the following works would be used: 
1.Descriptive Statistics
2.Assessing a variable for normality
3.Explain Type I and Type II errors
4.Explain Statistical Power and Statistical Significance
5.Correlation
6.Difference

For using this project the real data from Portuguese Secondary School Student Performance was used.
Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7. (https://repositorium.sdum.uminho.pt/bitstream/1822/8024/1/student.pdf).




Each of the variables of interest were inspected. The numerical variables of interest would be inspected for normality by checking standardised scores for skewness and kurtosis as well as considering the percentage of standardised scores for the variables fell outside of expected boundaries and creating histograms and QQ plots. Decisions on skewness and kurtosis were guided by the advice of (George & Mallory, 2011) which categorizes a distribution as normal if the relevant standardised scores for skewness and kurtosis fall in the range +/- 2 and the advice of (Field, Miles & Field, 2012) which categorizes a distribution as normal if 95% of the standardised scores for the variable fall within the bounds of +/-3.29 for a dataset larger than 80 cases. Summary statistics of various nominal variables were also identified for analysis.

## Normality
```{r, warning=FALSE, fig.width=10, fig.align='center'}
# Selecting the grades variable
grades <- student_data %>% select(pG1,pG2,pG3,mG1,mG2,mG3)

# create a custom variable called previous grades for Portuguese language(pG12)
pG12<-round((grades$pG1+grades$pG2)/2,0)

# create a custom variable called previous grade for Math(mG12)
mG12<-round((grades$mG1+grades$mG2)/2,0)

# add pG12 and mG12 variable to data as new column 
grades <- grades %>% 
  mutate(pG12,mG12) %>%
  select(-c(pG1,pG2,mG1,mG2))


# to get summary statistics for grades
describe(grades,omit = TRUE, IQR = TRUE)

# Distribution of grades variable with visualization
g1<-ggplot(grades,aes(x=pG12))+labs(x='PG12', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG1,na.rm = TRUE),sd=sd(q_var$pG1,na.rm = TRUE)))+
  ggtitle('Distribution Average previous Grade in Portuguese languag')+
  theme(plot.title = element_text(size=10))

g3<-ggplot(grades,aes(x=pG3))+labs(x='PG3', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG3,na.rm = TRUE),sd=sd(q_var$pG3,na.rm = TRUE)))+
  ggtitle('Distribution Final Grade in Portuguese languag')+
  theme(plot.title = element_text(size=10))

g4<-ggplot(grades,aes(x=mG12))+labs(x='MG12', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG1,na.rm = TRUE),sd=sd(q_var$mG1,na.rm = TRUE)))+
  ggtitle('Distribution Average previous Grade in Math')+
  theme(plot.title = element_text(size=10))

g6<-ggplot(grades,aes(x=mG3))+labs(x='MG3', y='Density')+
  geom_histogram(binwidth = 1,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG3,na.rm = TRUE),sd=sd(q_var$mG3,na.rm = TRUE)))+
  ggtitle('Distribution Final Grade in Math')+
  theme(plot.title = element_text(size=10))

grid.arrange(g1,g4,g3,g6,nrow=2)

# Create QQ Plot for previous grades for Portuguese language(pG12)
qqnorm(grades$pG12, main = "QQplot previous grades for Portuguese language(pG12)")
qqline(grades$pG12,col=2) # show line on the plot

# Create QQ Plot for previous grades for Portuguese language(pG12)
qqnorm(grades$mG12, main = "QQplot previous grade for Math(mG12)")
qqline(grades$mG12,col=2) # show line on the plot

# Create QQ Plot for final grades for Portuguese language(pG3)
qqnorm(grades$pG3, main = "QQplot final grades for Portuguese language(pG3)")
qqline(grades$pG3,col=2) # show line on the plot

# Create QQ Plot for final grades for Math(mG3)
qqnorm(grades$mG3, main = "QQplot final grades for Math(mG3)")
qqline(grades$mG3,col=2) # show line on the plot


```
## Report of Normality befor transformation
Students with no grades is a big problem in database because some of the have the grades for previous grade and don't have any grades for final grades and vice versa.Therefor for next part just the grades bigger than zero would be considered.



## Normality After Transformation
```{r, warning=FALSE, fig.width=10, fig.align='center'}
# select all grates than are biger than 0
ngrades <- grades %>% filter(pG12>0,pG3>0,mG3>0,mG12>0)

ngrades_d<-describe(ngrades,omit = TRUE, IQR = TRUE)
ngrades_s<-pastecs::stat.desc(ngrades, basic = FALSE)
ngrades_st <- ngrades_s  %>% transpose()
rownames(ngrades_st) <- colnames(ngrades_s)
colnames(ngrades_st) <- rownames(ngrades_s)

stdskew <- list()
stdkurt <- list()
gt196 <- list()
gt329 <- list()

for (n in 1:4) {
  v <- row.names.data.frame(ngrades_st)[n]

  skew        <- semTools::skew(ngrades[[v]])
  kurt        <- semTools::kurtosis(ngrades[[v]])
  stdskew[[v]] <- skew[1] / skew[2]
  stdkurt[[v]] <- kurt[1] / kurt[2]
  zscore        <- abs(scale(ngrades[[v]]))
  gt196[[v]]   <- FSA::perc(as.numeric(zscore), 1.96, "gt")
  gt329[[v]]   <- FSA::perc(as.numeric(zscore), 3.29, "gt")

}

ngrades_st$skew <- skew
ngrades_st$kurt <- kurt
ngrades_st$std.skew <- stdskew
ngrades_st$std.kurt <- stdkurt
ngrades_st$gt.196 <- gt196
ngrades_st$gt.329 <- gt329
ngrades_st

# Distribution of grades variable with visualization After Transformation
g1<-ggplot(ngrades,aes(x=pG12))+labs(x='PG12', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG1,na.rm = TRUE),sd=sd(q_var$pG1,na.rm = TRUE)))+
  ggtitle('Distribution Average previous Grade in Portuguese languag')+
  theme(plot.title = element_text(size=10))

g3<-ggplot(ngrades,aes(x=pG3))+labs(x='PG3', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$pG3,na.rm = TRUE),sd=sd(q_var$pG3,na.rm = TRUE)))+
  ggtitle('Distribution Final Grade in Portuguese languag')+
  theme(plot.title = element_text(size=10))

g4<-ggplot(ngrades,aes(x=mG12))+labs(x='MG12', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG1,na.rm = TRUE),sd=sd(q_var$mG1,na.rm = TRUE)))+
  ggtitle('Distribution Average previous Grade in Math')+
  theme(plot.title = element_text(size=10))

g6<-ggplot(grades,aes(x=mG3))+labs(x='MG3', y='Density')+
  geom_histogram(binwidth = 2,colour='black',aes(y=..density..,fill=..count..))+
  scale_fill_gradient("Count",low="#132B43", high="#56B1F7")+
  stat_function(fun = dnorm,color="red",args = list(mean=mean(q_var$mG3,na.rm = TRUE),sd=sd(q_var$mG3,na.rm = TRUE)))+
  ggtitle('Distribution Final Grade in Math')+
  theme(plot.title = element_text(size=10))


grid.arrange(g1,g4,g3,g6,nrow=2)

# Checking Normality by QQ-Plot After Transformation
# Create QQ Plot for previous grades for Portuguese language(pG12)
qqnorm(ngrades$pG12, main = "QQplot previous grades for Portuguese language(pG12)")
qqline(ngrades$pG12,col=2) # show line on the plot

# Create QQ Plot for previous grades for Portuguese language(pG12)
qqnorm(ngrades$mG12, main = "QQplot previous grade for Math(mG12)")
qqline(ngrades$mG12,col=2) # show line on the plot

# Create QQ Plot for final grades for Portuguese language(pG3)
qqnorm(ngrades$pG3, main = "QQplot final grades for Portuguese language(pG3)")
qqline(ngrades$pG3,col=2) # show line on the plot

# Create QQ Plot for final grades for Math(mG3)
qqnorm(ngrades$mG3, main = "QQplot final grades for Math(mG3)")
qqline(ngrades$mG3,col=2) # show line on the plot

# One-sample K-S Normality Test (performed if samples > 50)
ks.test(ngrades$pG12,"pnorm",mean=mean(ngrades$pG12,na.rm = TRUE),sd=sd(ngrades$pG12,na.rm = TRUE))
ks.test(ngrades$mG12,"pnorm",mean=mean(ngrades$mG12,na.rm = TRUE),sd=sd(ngrades$mG12,na.rm = TRUE))
ks.test(ngrades$pG3,"pnorm",mean=mean(ngrades$pG3,na.rm = TRUE),sd=sd(ngrades$pG3,na.rm = TRUE))
ks.test(ngrades$mG3,"pnorm",mean=mean(ngrades$mG3,na.rm = TRUE),sd=sd(ngrades$mG3,na.rm = TRUE))

```
## Report Of Normality

Average previous grade is represented by a Numerical variable  which has been calculated by average of variables that namely (pG1,pG2) for Average previous grade of Portuguese(pG12) and (mG1,mG2) for Average previous grade of Math(mG12).
Inspection of the standardized scores:
Average previews grade of Portuguese(pG12) for skewness (*skewness* = 2.1) and kurtosis (*kurtosis* = -1.08) reveal that just kurtosis values fall within the standardised score range of -2 and 2. This implies that  kurtosis appear to be normal. On further inspection using plots like histogram and normality plot, the distribution appears normal. On inspection of the count of outliers, we found only 0% standardised scores were outside the acceptable range of [-3.29, +3.29] which implies none of the values is outside the 95% Confidence Interval. Further analysis of the variable with One-Sample Kolmogorov-Smirnov Normality Test, we found the *p-value* = 2.698e-07 (which is less than standard value of .05) which  can reject the null hypothesis that the data is not normally distributed. Conclusively, based on all the tests, we can say that the data for Average previews grade of Portuguese(pG12) variable would be be treated as a normal within this analysis (*M* = `r mean(ngrades$pG12, na.rm=TRUE)`, *SD* = `r sd(ngrades$pG12, na.rm=TRUE)`, *N* = `r length(ngrades$pG12)-sum(is.na(ngrades$pG12))`).

Average previews grade of Math(mG12) for skewness (*skewness* = 1.74) and kurtosis (*kurtosis* =-2.52) reveal that just skewness values fall within the standardised score range of -2 and 2. This implies that  skewness appear to be normal. On further inspection using plots like histogram and normality plot, the distribution appears normal. On inspection of the count of outliers, we found only 0% standardised scores were outside the acceptable range of [-3.29, +3.29] which implies none of the values is outside the 95% Confidence Interval. Further analysis of the variable with One-Sample Kolmogorov-Smirnov Normality Test, we found the *p-value* = 0.00013 (which is less than standard value of .05) which  reject the null hypothesis that the data is not normally distributed. Conclusively, based on all the tests, we can say that the data for Average previews grade of Math(mG12) variable would be be treated as a normal within this analysis (*M* = `r mean(ngrades$mG12, na.rm=TRUE)`, *SD* = `r sd(ngrades$mG12, na.rm=TRUE)`, *N* = `r length(ngrades$mG12)-sum(is.na(ngrades$mG12))`).

Final grade of Portuguese(pG3) for skewness (*skewness* = -1.43) and kurtosis (*kurtosis* = 3.06) reveal that just skewness values fall within the standardised score range of -2 and 2. This implies that  skewness appear to be normal. On further inspection using plots like histogram and normality plot, the distribution appears normal. On inspection of the count of outliers, we found only  0.29% standardised scores were outside the acceptable range of [-3.29, +3.29] which implies none of the values is outside the 95% Confidence Interval. Further analysis of the variable with One-Sample Kolmogorov-Smirnov Normality Test, we found the *p-value* =  0.00124 (which is less than standard value of .05) which  can reject the null hypothesis that the data is not normally distributed. Conclusively, based on all the tests, we can say that the data for Final grade of Portuguese(pG3) variable would be be treated as a normal within this analysis (*M* = `r mean(ngrades$pG3, na.rm=TRUE)`, *SD* = `r sd(ngrades$pG3, na.rm=TRUE)`, *N* = `r length(ngrades$pG3)-sum(is.na(ngrades$pG3))`).

Final grade of Math(mG) for skewness (*skewness* = 1.63) and kurtosis (*kurtosis* =--1.72,) reveal that both skewness and kurtosis values fall within the standardised score range of -2 and 2. This implies that  both skewness and kurtosis appear to be normal. On further inspection using plots like histogram and normality plot, the distribution appears normal. On inspection of the count of outliers, we found only 0% standardised scores were outside the acceptable range of [-3.29, +3.29] which implies none of the values is outside the 95% Confidence Interval. Further analysis of the variable with One-Sample Kolmogorov-Smirnov Normality Test, we found the *p-value* = 0.0002 (which is less than standard value of .05) which  reject the null hypothesis that the data is not normally distributed. Conclusively, based on all the tests, we can say that the data for Final grade of Math(mG12) variable would be be treated as a normal within this analysis (*M* = `r mean(ngrades$mG, na.rm=TRUE)`, *SD* = `r sd(ngrades$mG, na.rm=TRUE)`, *N* = `r length(ngrades$mG)-sum(is.na(ngrades$mG))`).



Hypothesis 1

H0: there is no relationship between students Average previous grades and final grade for Portuguese major.
H1: there is a relationship between students Average previous grades and final grade for Portuguese major.

n alpha level .05 was adopted and Cohen's conventions on effect size (coefficient=r) was adopted for Pearson Correlation Test. According to Cohen, the effect size is low if the value of r varies around .1, medium if r varies around .3, and large if r varies more than .5 (Field, Miles & Field, 2012).

```{r, warning=FALSE, fig.width=10, fig.align='center'}
# Pearson Correlation test for normal data distribution
cor.test(ngrades$pG12,ngrades$pG3,method = 'pearson')
# (r=0.89, n=338, p= 2.2e-16)

# Create scatter plot
s1<-ggplot()+aes(x=ngrades$pG12,y=ngrades$pG3)
s1<-s1+geom_point()+geom_smooth(method = 'lm',colour='red',se = FALSE)+
  labs(x='Average of Previous portuguese Grade',y='Final portuguese Grade')+ 
  ggtitle('Correlation between Average of Previous portuguese Grade and Final Portuguese Grade')+
  theme(plot.title = element_text(size=12))

s1

# calculate covariance
na.omit(ngrades$pG12,ngrades$pG3)
cov_p <- cov(ngrades$pG12,ngrades$pG3,method = "pearson")
cov_p
```
## Report of Correlation Analysis

The relationship between Final Portuguese Grades  and Average previous Portuguese Grades (that comes from average of pG1 and pG2) was investigated using a Pearson Correlation Test. A higher than moderate positive correlation was found with an effect size of 0.89 and direction of relationship being positive between both variables (*r* = 0.89, *n* = 338, *p* = 2.2e-16). The scatter plot  also validates this correlation. Final Portuguese Grades and and Average previous Portuguese Grades have shared variance ($(r^2)*100$) of 79% and cov_p of `r cov_p` . From the correlation test, we found the p-value (2.2e-16) less than the significance level of .05 which implies that this is sufficient evidence to reject the null hypothesis that this correlation between these variables is due to random chance. Hence, based on the test and scatter plot, we conclude with enough evidence that there is more than moderate positive correlation between Final Portuguese Grades  and Average previous Portuguese Grades . And we can conclude with confidence that this relationship is genuine and not due to random chance.


Hypothesis 2

H0: there is no relationship between students Average previous grades and final grade for Math major.
H1: there is a relationship between students Average previous grades and final grade for Math major.
```{r, warning=FALSE, fig.width=10, fig.align='center'}
# Pearson Correlation test for normal data distribution
cor.test(ngrades$mG12,ngrades$mG3,method = 'pearson')
# (r=-0.94, n=338, p= 2.2e-16)

# Create scatter plot
s2<-ggplot()+aes(x=ngrades$mG12,y=ngrades$mG3)
s2<-s2+geom_point()+geom_smooth(method = 'lm',colour='red',se = FALSE)+
  labs(x='Average of Previous Math Grade',y='Final Math Grade')+ 
  ggtitle('Correlation between Average of Previous Math Grade and Math Final Math Grade ')+
  theme(plot.title = element_text(size=12))
s2

# calculate covariance
na.omit(ngrades$mG12,ngrades$mG3)
cov_m <- cov(ngrades$mG12,ngrades$mG3,method = "pearson")
cov_m

```

## Report of Correlation Analysis

The relationship between Final Math Grades  and Average previous Math Grades (that comes from average of mG1 and mG2) was investigated using a Pearson Correlation Test. A higher than moderate positive correlation was found with an effect size of 0.89 and direction of relationship being positive between both variables (*r* = 0.94, *n* = 338, *p* = 2.2e-16). The scatter plot  also validates this correlation. Final Math Grades and and Average previous Math Grades have shared variance ($(r^2)*100$) of 88% and cov_m of `r cov_m` . From the correlation test, we found the p-value (2.2e-16) less than the significance level of .05 which implies that this is sufficient evidence to reject the null hypothesis that this correlation between these variables is due to random chance. Hence, based on the test and scatter plot, we conclude with enough evidence that there is more than moderate positive correlation between Final Math Grades  and Average previous Math Grades . And we can conclude with confidence that this relationship is genuine and not due to random chance.


Hypothesis 3

H0: There is no difference in Final Portuguese Grade of students  and their study time.
HA: There is a difference in Final Portuguese Grade of students and their study time.

An alpha level .05 was adopted and Cohen's conventions on effect size (eta-squared) was adopted for ANOVA Test. According to Cohen, the effect size is low if the value of eta-squared varies around .01, medium if eta-squared varies around .06, and large if eta-squared varies more than .14 (Field, Miles & Field, 2012).

```{r, warning=FALSE, fig.width=10, fig.align='center'}
# using Parametric Test
student_data <- student_data %>% filter(pG3>0,mG3>0)

# Descriptive statistics by study time status
describeBy(student_data$pG3,student_data$studytime.p)

# check mean for each study time status group
mean_study_p<-round(tapply(student_data$pG3,student_data$studytime.p,FUN =mean),digits = 2)
mean_study_p

# The plot shows how mean values of Final Grades change with study time status for Final Portuguese Grade
plotmeans(student_data$pG3~student_data$studytime.p,digits = 2,
          ccol = 'red',mean.labels = TRUE,xlab = 'Study Time Status',ylab = 'Final Portuguese Grade',
          main='Plot of Final Portuguese Grade Means by Study Time Status')

# Create boxplot
boxplot(student_data$pG3~student_data$studytime.p,
        main='Final Portuguese Grade Means by Study Time Status (mean is by black dot)',
        xlab = 'Study Time Status',ylab = 'Final Portuguese Grade',col=c("#132B43","#56B1F7","#56B1F7","#132B43"))
points(mean_study_p,col='black',pch=18)

bartlett.test(student_data$pG3,student_data$studytime.p) 
# p-value= 0.44 > .05, 
# null hypothesis of the test is accepted i.e. the variance of different groups can be assumed to be equal.

# Perform ANOVA test
aov_study_p<-aov(student_data$pG3~student_data$studytime.p)
summary(aov_study_p)

# Calculate effect size
etaSquared(aov_study_p)

```
## Report of Difference Analysis

A Bartlett's test was conducted and indicated equality of variance for Final Portuguese Grades for all groups of various Rate of Study Time (*K-squared* = 2.6735, *P* = 0.4447). A one-way between-groups analysis of variance was conducted to explore Final Math Grade for respondents having Rate of Study Time. Participants were divided into  groups according to their Rate of Study Time (Group 1 :  <2 hours; Group 2 : 2 to 5 hours; Group 3 :5 to 10 hours; Group 4 :>10 hours). There was statistically  difference level in Final Math Grade mean scores for respondents of different Rate of Study Time (*F(5, 260)* = 24.87, *p* = 9.82e-07) p-value is very small. The effect size, calculated using eta squared was 0.06 which implies there is a strong standardized mean difference for different groups. The test results indicate there is evidence to support rejecting the null hypothesis that there is a strong difference in Final Portuguese Grades for respondents of different Rate of Study Time.


Hypothesis 4

H0: There is no difference in Final Math Grade of students and their study time.
HA: There is a difference in Final Math Grade of students and their study time.

An alpha level .05 was adopted and Cohen's conventions on effect size (eta-squared) was adopted for ANOVA Test. According to Cohen, the effect size is low if the value of eta-squared varies around .01, medium if eta-squared varies around .06, and large if eta-squared varies more than .14 (Field, Miles & Field, 2012).

```{r, warning=FALSE, fig.width=10, fig.align='center'}
# Since the Tperformance data is normally distributed, we are using Parametric Test
student_data <- student_data %>% filter(pG3>0,mG3>0)

as.factor(student_data$studytime.m)
# Descriptive statistics by study time status
describeBy(student_data$mG3,student_data$studytime.m)

# check mean for each study time status group
mean_study_m<-round(tapply(student_data$mG3,student_data$studytime.m,FUN =mean),digits = 2)
mean_study_m


# The plot shows how mean values of Final Math Grades change with study time status 
plotmeans(student_data$mG3~student_data$studytime.m,digits = 2,
          ccol = 'red',mean.labels = TRUE,xlab = 'Study Time Status',ylab = 'Final Math Grade',
          main='Plot of Final Math Grade Means by Study Time Status')

# Create boxplot
boxplot(student_data$mG3~student_data$studytime.m,
        main='Final Math Grade Means by Study Time Status (mean is by black dot)',
        xlab = 'Study Time Status Group',ylab = 'Final Math Grade',col=c("#132B43","#56B1F7","#56B1F7","#132B43"))
points(mean_study_p,col='black',pch=18)

# Bartlett test of homogeneity of variances
bartlett.test(student_data$mG3,student_data$studytime.m) 
# p-value= 0.79 > .05, 
# null hypothesis of the test is accepted i.e. the variance of different groups can be assumed to be equal.


# Perform ANOVA test
aov_study_m<-aov(student_data$mG3~student_data$studytime.m)
summary(aov_study_m)

# Since the value of F-statistic=  3.897 > 1 and p-value= 0.0492 < 0.05, this implies that the variation among the sample means is higher than the variation within groups, so the mean values for different groups are  different but this diffrenece is not much because the p-value is near to 0.5. Thus, we can reject the null hypothesis of the test that mean of values for different groups are equal and thereby accept the alternate hypothesis.
# There is statistically significant difference in mean values of Final MaTH Grade for respondents of different Rate of Study Time.

# Calculate effect size
etaSquared(aov_study_m)


```

## Report of Difference Analysis

A Bartlett's test was conducted and indicated equality of variance for Final Math Grades for all groups of various Rate of Study Time (*K-squared* = 1.008, *P* = 0.7993). A one-way between-groups analysis of variance was conducted to explore Final Math Grade for respondents having Rate of Study Time. Participants were divided into  groups according to their Rate of Study Time (Group 1 :  <2 hours; Group 2 : 2 to 5 hours; Group 3 :5 to 10 hours; Group 4 :>10 hours). There was statistically  difference level in Final Math Grade mean scores for respondents of different Rate of Study Time (*F(5, 260)* = 3.897, *p* = 0.0492) p-value is smaller than 0.5 but it is near the 0.5. The effect size, calculated using eta squared was 0.01 which implies there is no near strong standardized mean difference for different groups. The test results indicate there is evidence to support accepting the null hypothesis that there is difference in Final Math Grades for respondents of different Rate of Study Time but this difference is not strong.

